What is happening on Twitter? A framework for student research projects with tweets

### 2020-09-11 Rmd/jse-special-issue-2020-october.Rmd
added rtweet reference at first mention
added mention of curse words
fixed "We also note" issue

We include here instructions for creating a tweet collection. First, we created a 
new account on Twitter. With these user credentials, we used the 
R package `rtweet` to query the API. We used the linux `crontab` software 
to repeatedly execute R code to submit API queries. Each 
query lasted five minutes and produced a text file of approximately 130 MB. We timed the API queries so 
that there was no time lag between queries. We stored tweets resulting from API queries in their native JSON format.

@@

We identified the top ten most probable terms for each of ten topics in our models 
(Figures \@ref(fig:may24), \@ref(fig:may25), \@ref(fig:may26)). 
We plotted the within-topic word probabilities as bar graphs. We see that topic-specific 
word probabilities seldom exceed 0.05. We also note that some words are heavily weighted in 
multiple topics. This observation complicates semantic topic interpretation.

@@

Assigning meaning to topics is an active research area [@chang2009reading]. Since
our interest is in the transient appearance of a new topic, we don't attempt to
assign meaning to every topic in our models. We see that topic 7 from May 25 has several 
words that suggest Memorial Day: memorial, remember, honor, country. A similar topic is not 
seen on May 24 or May 26. Some topics persist, with distinct word probabilities, 
across the three days. For example, we see that President Trump features prominently in all 
three models. We also note, on May 26, topic 10, which reflects discussion of the Amy 
Cooper Central Park incident (https://www.nytimes.com/2020/05/26/nyregion/amy-cooper-dog-central-park.html).

We also note that the murder of George Floyd occurred on May 25. Our last 
examined time period, from 12:00 pm to 12:05 pm (Eastern USA time zone) on May 26, occurred after Floyd's murder, 
yet we didn't detect this event in our ten-topic LDA model. Several considerations 
may account for this. While outrage at the murder eventually spread worldwide, there may have been few 
Floyd-related tweets during our collection time on May 26, less than 24 hours after the murder and video release. 
Had we extended our analysis to May 27 and beyond, we may have identified George Floyd-related topics.

>>>>

We include here instructions for creating a tweet collection. First, we created a 
new account on Twitter. With these user credentials, we used the 
R package `rtweet` to query the API [@rtweet-package]. We used the linux `crontab` software 
to repeatedly execute R code to submit API queries. Each 
query lasted five minutes and produced a text file of approximately 130 MB. We timed the API queries so 
that there was no time lag between queries. We stored tweets resulting from API queries in their native JSON format.

@@

We identified the top ten most probable terms for each of ten topics in our models 
(Figures \@ref(fig:may24), \@ref(fig:may25), \@ref(fig:may26)). 
We plotted the within-topic word probabilities as bar graphs. We see that topic-specific 
word probabilities seldom exceed 0.05. We also note that some words are heavily weighted in 
multiple topics. This observation complicates semantic topic interpretation. We 
also caution that the results display expletives (that appeared on Twitter) and may 
not be "safe for work".

@@

Assigning meaning to topics is an active research area [@chang2009reading]. Since
our interest is in the transient appearance of a new topic, we don't attempt to
assign meaning to every topic in our models. We see that topic 7 from May 25 has several 
words that suggest Memorial Day: memorial, remember, honor, country. A similar topic is not 
seen on May 24 or May 26. Some topics persist, with distinct word probabilities, 
across the three days. For example, we see that President Trump features prominently in all 
three models. On May 26, topic 10 reflects discussion of the Amy 
Cooper Central Park incident (https://www.nytimes.com/2020/05/26/nyregion/amy-cooper-dog-central-park.html).

The murder of George Floyd occurred on May 25. Our last 
examined time period, from 12:00 pm to 12:05 pm (Eastern USA time zone) on May 26, occurred after Floyd's murder, 
yet we didn't detect this event in our ten-topic LDA model. Several considerations 
may account for this. While outrage at the murder eventually spread worldwide, there may have been few 
Floyd-related tweets during our collection time on May 26, less than 24 hours after the murder and video release. 
Had we extended our analysis to May 27 and beyond, we may have identified George Floyd-related topics.
